---
title: "POLI3148 In-Class Exercise 2"
author: "Bernadette Marie Coronel La'O 3035957329"
date: "2023-10-05"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
d <- read_csv("/Users/berni/Desktop/POLI3148 PORTFOLIO/_DataPublic_/vdem/1984_2022/vdem_1984_2022_external.csv", show_col_types = FALSE)
```

Link:

### 1. Codebook lookup. Look up the codebook, answer the following questions:

1.  What indicators regarding the quality of education are available in the V-Dem datasets?

    ```{r}
    d |> select(e_peaveduc, e_peedgini)
    ```

    2.  What are the data's coverage (i.e., for which countries and years do we have data?)

    ```{r}
    d |> select(country_name, year) |> distinct()
    ```

    3.  What are their sources? Provide the link to at least 1 source.

    ```{r}
    # Clio Infra (clio-infra.eu), drawing on Mitchell (1998a, 1998b, 1998c), United States Census Bureau (2021), UNESCO, Földvári and van Leeuwen (2014), Leeuwen, van Leeuwen- Li, Földvári (2011), Leeuwen, van Leeuwen-Li, Földvári (2012a), Leeuwen, van Leeuwen-Li, Földvári (2012b), Didenko, Foldvari, van Leeuwen (2012)
    ```

    ### **2. Subset by columns**

    1.  Create a dataset containing only the country-year identifiers and indicators of education quality.

    ```{r}
    d_eq <- d |>
      select(country_name, year, e_peaveduc, e_peedgini)
    d_eq
    ```

    2.  Rename the columns of education quality to make them informative.

    ```{r}
    d_eqr <- d_eq |>
      rename("Country" = "country_name", "Year" = "year", "Education" = "e_peaveduc", "Inequality" = "e_peedgini")
    d_eqr
    ```

    ### **3. Subset by rows**

    1.  List 5 countries-years that have the highest education level among its population.

    ```{r}
    d_eqr |>
      slice_max(order_by = Education, n = 5)
    ```

    2.  List 5 countries-years that suffer from the most severe inequality in education.

    ```{r}
    d_eqr |>
      slice_max(order_by = Inequality, n = 5)
    ```

    ### **4. Summarize the data**

    1.  Check data availability: For which countries and years are the indicators of education quality available?

    ```{r}
    d_eqr |>
      group_by(Country) |>
      mutate(eq_missing = as.numeric(is.na(Education))) |>
      arrange(-eq_missing)
    ```

    2.  Create average level of education quality from 1984 to 2022

    ```{r}
    d_eqr |>
      group_by(Country) |>
      summarise(eq_average = mean(Education, na.rm = TRUE)) |>
      arrange(-eq_average)
    ```

    3.  Create change of education quality from 1984 to 2022

    ```{r}
    d_eqr |>
      group_by(Country) |>
      arrange(Year) |>
      mutate(eq_yoy_change = Education - lag(Education, n=1)) |>
      ungroup() |>
      arrange(Country, Year) |>
      arrange(-eq_yoy_change)
    ```

    4.  Examine the data and *briefly* discuss: Which countries perform the best and the worst in terms of education quality in the past four decades?

    ```{r}
    # Germany, Australia, and the UK have the highest education averages throughout 1984-2022.
    # Botswana, Singapore, and Thailand have had the most improvements.

    # Burkina Faso, Niger, and Mali have the lowest education averages throughout 1984-2022.
    # Namibia, Russia, and Sweden have had the least improvements.
    ```
